Phenology and Mycorrhizae Manuscript

### 2023-02-09 manuscriptCode_phenologyAndMycorrhizae.Rmd
Added scripts to compute weekly met data based off of dates of phenophases transitions

#Load Functions and Variables
```{r}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
downloadData=FALSE #Change to true if you need to download data 
```

Download NEON Data

#NEON Phenology Observations 
```{r,eval=FALSE}
source('combineNEONdata_PhenoObs.R')
dataName="NEON_PhenologyObservations"
NEON_ID='DP1.10055.001'

@@

#Single Air Temperature At Various Heights 
```{r,eval=FALSE}
dataName="NEON_SingleAirTemperature"
NEON_ID='DP1.00002.001'
selectColumns <- c('horizontalPosition','verticalPosition','tempSingleMean','tempSingleExpUncert','finalQF')
inFileName <- "SAAT_30min.csv"
varName <- 'tempSingleMean'
if(downloadData){
downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
for(funName in c("mean","max","min")){
  calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=mean,funName=funName)
  #Calculate LRs
  gapFillFromDayMet_verticalProfiles(dataName=dataName,dataPath=dataPath,varName=varName,funName=funName)
}

@@

#NEON Precipitation
```{r,eval=FALSE}
dataName="NEON_PrecipitationData"
NEON_ID='DP1.00006.001'
selectColumns <- c('precipBulk','precipExpUncert','precipQF')
inFileName <- c("PRIPRE_30min.csv","SECPRE_30min.csv")
varName <- "precipBulk"
if(downloadData){
downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=sum)

gapFillFromDaymet(dataName=dataName,dataPath=dataPath,varName=varName)

@@

#Download Photosynthetically active radiation (PAR)
```{r}
dataName="NEON_PAR"
NEON_ID='DP1.00024.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','PARMean','PARMinimum','PARMaximum','PARVariance','PARFinalQF')
inFileName <- "PARPAR_30min.csv"
varName <- "PARMean"

if(downloadData){
downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
funType=mean
funName="mean"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=min
funName="min"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=max
funName="max"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

@@

#Download Relative Humidity
```{r}
dataName="NEON_relativeHumidity"
NEON_ID='DP1.00098.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','RHMean','RHMinimum','RHMaximum','RHVariance','dewTempMean',
                   'RHFinalQF','dewTempFinalQF')
inFileName <- "RH_30min.csv"
varName <- "RHMean"

@@

funType=mean
funName="mean"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=min
funName="min"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=max
funName="max"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

#Download Windspeed
```{r}
dataName="NEON_Windspeed"
NEON_ID='DP1.00001.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','windSpeedMean','windSpeedMinimum','windSpeedMaximum','windSpeedFinalQF')
inFileName <- "2DWSD_30min.csv"
varName <- "windSpeedMean"

if(downloadData){
downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
funType=mean
funName="mean"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=min
funName="min"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=max
funName="max"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

#Download Soil Temp
```{r}
dataName="NEON_SoilTemp"
NEON_ID='DP1.00041.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

@@

funType=mean
funName="mean"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=min
funName="min"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

funType=max
funName="max"
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType,funName=funName)

#Download NEON Soil Microbe Data
```{r,eval=FALSE}
dataName="NEON_SoilMicrobeComposition"
NEON_ID='DP1.10081.001'
#selectColumns <-
inFileName <- c("mcc_soilSeqVariantMetadata_16S.csv","mcc_soilSeqVariantMetadata_ITS.csv")
#varName <- 
if(downloadData){
downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID) #Note: I had to modify this function to remove the indices for sites that didn't seem to have data. 
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
```

#Download NEON Soil Physical and Chemical Properties
```{r,eval=FALSE}
dataName <- "NEON_soilProperties"
NEON_ID <- "DP1.10086.001"
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
if(downloadData){
downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}

@@

@@

#NEON Root Characteristics
```{r, eval=FALSE}
dataName="NEON_Roots"
NEON_ID='DP1.10067.001'

@@

#NEON Plant Structural Characteristics
```{r}
dataName="NEON_vegetation_structure"
NEON_ID="DP1.10098.001"

@@

#NEON Litterfall
```{r}
dataName="NEON_litterfall" #No: 2, 7,8, 11, 17, 18, 20, 23, 25, 27, 38, 41, 45
NEON_ID="DP1.10033.001"
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

@@

#NEON Foliar Traits
```{r}
dataName="NEON_plantFoliarTraits"
NEON_ID="DP1.10026.001"

@@

outFileName <- paste0(dataName,"ALLdata.csv")
write.csv(file=paste0(dataPath,outFileName),allDat,row.names = FALSE,quote=FALSE)

@@

#NEON PhenoCam Observations
```{r,eval=FALSE}
source('determinePhenoCamURLs.R')
source('downloadPhenoCamAndCalculateTransitions.R') #Still need to change to have it save it somewhere

>>>>

#Load Functions and Variables
```{r}
source('sharedVariables.R')
source('NEON_Data_DownloadAndProcess.R')
source('computeWeeklyMetData.R')
downloadData=FALSE #Change to true if you need to download data 
```

#Download and Process Individual NEON Data Products

##NEON Phenology Observations 
```{r,eval=FALSE}
source('combineNEONdata_PhenoObs.R')
dataName="NEON_PhenologyObservations"
NEON_ID='DP1.10055.001'

@@

##NEON Single Air Temperature At Various Heights 
```{r,eval=FALSE}
dataName="NEON_SingleAirTemperature"
NEON_ID='DP1.00002.001'
selectColumns <- c('horizontalPosition','verticalPosition','tempSingleMean','tempSingleExpUncert','finalQF')
inFileName <- "SAAT_30min.csv"
varName <- 'tempSingleMean'
if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)

for(f in seq_along(funTypes)){
  calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funTypes[f],funName=funNames[f])
  #Calculate LRs
  # gapFillFromDayMet_verticalProfiles(dataName=dataName,dataPath=dataPath,varName=varName,funName=funName)
} #****Need to rerun

@@

##NEON Precipitation
```{r,eval=FALSE}
dataName="NEON_PrecipitationData"
NEON_ID='DP1.00006.001'
selectColumns <- c('precipBulk','precipExpUncert','precipQF')
inFileName <- c("PRIPRE_30min.csv","SECPRE_30min.csv")
varName <- "precipBulk"
if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=sum)

#gapFillFromDaymet(dataName=dataName,dataPath=dataPath,varName=varName)

@@

##NEON Photosynthetically active radiation (PAR)
```{r}
dataName="NEON_PAR"
NEON_ID='DP1.00024.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','PARMean','PARMinimum','PARMaximum','PARVariance','PARFinalQF')
inFileName <- "PARPAR_30min.csv"
varName <- "PARMean"

if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)

for(f in seq_along(funTypes)){
  calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funTypes[f],funName=funNames[f])
}

@@

##NEON Relative Humidity
```{r}
dataName="NEON_relativeHumidity"
NEON_ID='DP1.00098.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','RHMean','RHMinimum','RHMaximum','RHVariance','dewTempMean',
                   'RHFinalQF','dewTempFinalQF')
inFileName <- "RH_30min.csv"
varName <- "RHMean"

@@

for(f in seq_along(funTypes)){
  calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funType[f],funName=funName[f])
}
```

##NEON Windspeed
```{r}
dataName="NEON_Windspeed"
NEON_ID='DP1.00001.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
selectColumns <- c('verticalPosition','windSpeedMean','windSpeedMinimum','windSpeedMaximum','windSpeedFinalQF')
inFileName <- "2DWSD_30min.csv"
varName <- "windSpeedMean"

if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)

for(f in seq_along(funTypes)){
  calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funTypes[f],funName=funNames[f])
}
```

##NEON Soil Temp
```{r}
dataName="NEON_SoilTemp"
NEON_ID='DP1.00041.001'
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

@@

for(f in seq_along(funTypes)){
  calculateDailyWeather(dataName=dataName,dataPath=dataPath,varName=varName,funType=funTypes[f],funName=funNames[f])
}
```

##NEON Soil Microbe Data
```{r,eval=FALSE}
dataName="NEON_SoilMicrobeComposition"
NEON_ID='DP1.10081.001'
#selectColumns <-
inFileName <- c("mcc_soilSeqVariantMetadata_16S.csv","mcc_soilSeqVariantMetadata_ITS.csv")
#varName <- 
if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID) #Note: I had to modify this function to remove the indices for sites that didn't seem to have data. 
}
combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath)
```

##NEON Soil Physical and Chemical Properties
```{r}
dataName <- "NEON_soilProperties"
NEON_ID <- "DP1.10086.001"
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}
if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID)
}

@@

allDat %>% filter(horizon!="")
allDat <- pivot_wider(allDat,names_from=6,values_from=7:11)

##Still need to run:
inFileName <- "sls_soilChemistry.csv"
selectColumns <- c('plotID','d15N','organicd13C','nitrogenPercent','organicCPercent',
                   'CNratio')
dat4 <- combineNEONdata(dataName=dataName,NEON_ID=NEON_ID,selectColumns=selectColumns,
                inFileName=inFileName,dataPath=dataPath,saveFile = FALSE)
dat4 <- dat4 %>% mutate(collectDate=format(as.POSIXct(collectDate,                                                  format='%Y-%m-%dT%H:%M'),format="%Y-%m-%d")) %>% 
  mutate(year=lubridate::year(collectDate)) %>% group_by(siteID,year) %>%
  summarise(d15N=mean(d15N),organicd13C=mean(organicd13C),
            nitrogenPercent=mean(nitrogenPercent),organicCPercent=mean(organicCPercent),
            CNratio=mean(CNratio)) 
allDat <- full_join(allDat,dat4,by=c('siteID','year'))

@@

##NEON Root Characteristics
```{r}
dataName="NEON_Roots"
NEON_ID='DP1.10067.001'

@@

##NEON Plant Structural Characteristics
```{r}
dataName="NEON_vegetation_structure"
NEON_ID="DP1.10098.001"

@@

##NEON Litterfall
```{r}
dataName="NEON_litterfall"
NEON_ID="DP1.10033.001"
if(!file.exists(paste0(dataPath,"/",dataName))){
  dir.create(paste0(dataPath,"/",dataName))
}

if(downloadData){
  downloadNEONdata(dataName=dataName,NEON_ID=NEON_ID,includedSeq=(1:47)[-c(2,7,8,11,17,18,20,23,25,27,38,41,45)])
}

@@

##NEON Foliar Traits
```{r}
dataName="NEON_plantFoliarTraits"
NEON_ID="DP1.10026.001"

@@

##NEON PhenoCam Observations
```{r,eval=FALSE}
source('determinePhenoCamURLs.R')
source('downloadPhenoCamAndCalculateTransitions.R') #Still need to change to have it save it somewhere
```

##ERA5 Met Data
```{r}
source('downloadERA5.R')
source('combineERA5data.R')
```

##Calculate NEON Photoperiods
```{r}
source('calculateNEON_photoperiods.R')

@@

#Combine NEON Data for each site
```{r}
